### 2020-05-07 phylota-overview.Rmd
comparison with other tools

# Phylota overview

Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
database [@sanderson2008phylota].

Since then, it has been cited 122 times for different reasons.

1. As an example study of very large and comprehensive phylogeny from minedDNA sequence data, (even if no phylogeny was published there, only the method to do so):
  - @beaulieu2012modeling
  -

1. As an example tool that mines GenBak data for phylogenetic reconstruction:
  - original publication of PHLAWD [@smith2009mega]
  -

1. When the software was actually used to construct a DNA data set for phylogenetic reconstruction of:
  - A 1000 tip phylogeny of the family of the nightshades [@sarkinen2013solanaceae]
  - A 56 tip phylogeny of crustacean zooplancton [@helmus2010communities] -- ecological study
  - A 63 tip phylogeny of the Salmonidae family [@crete2012salmonidae]
  - A 321 tip phylogeny of Testudines [@thomson2010sparse]

1. When the website was used to identify sequences and markers available in
GenBank for a particular group, In this cases, the dataset mining was performed with other tools, or not performed at all and just used for discussion:
  - A 812 tips phylogeny of the Order Chiroptera [@shi2015speciation] -- dataset constructed with PHLAWD
  - A 1276 tips phylogeny of the Fabaceae [@legume2013legume] -- dataset constructed by hand (I think??)
  - A review of dated phylogenies of fire-prone tropical savanna species from Brazil [@simon2012cerrado] -- just for discussion of the lack of markers available for these species on GenBank

1. Sometimes, it was cited by mistake:
    - In this 630 tip phylogeny of the Caryophyllaceae study [@greenberg2011caryophyllaceae] it might have been originally
    cited as an example of large phylogenies that reflect well supported relationships
    from previous smaller phylogenies. However, it was removed from the text but
    not from the final list of references. The DNA data set was constructed by hand most probably.