### 2020-05-08 sysbio_submission.Rmd

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'Physcraper: A Python package for continually updated gene trees'
tags:
  - Python
  - phylogenetics
  - update
  - species tree
  - Open Tree
authors:
  - name: Luna L. Sanchez Reyes
    orcid: 0000-0000-0000-0000
    affiliation: 1 # (Multiple affiliations must be quoted, like this: "1, 2")
  - name: Emily Jane McTavish
    orcid: 0000-0000-0000-0000
    affiliation: 1
affiliations:
 - name: University of California, Merced
   index: 1
citation_author: Sanchez-Reyes et. al.
date: 24 April 2020
year: 2020
bibliography: paper.bib
output:
  pdf_document:
    template: template.tex
csl: apa.csl
links-as-notes: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  eval = TRUE
)
```

```{r body, child="physcraper_ms.Rmd"}
```