### 2020-05-13 phylota-overview.Rmd
some more phylota examples

# Phylota overview

Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
database [@sanderson2008phylota].

Since then, it has been cited 122 times for different reasons.

1. As an example study of very large and comprehensive phylogeny from mined DNA sequence data, (even if no phylogeny was really published there, only the method to do so):
  - @beaulieu2012modeling
  -

1. As an example of a tool that mines GenBank data for phylogenetic reconstruction:
  - original publication of PHLAWD [@smith2009mega]
  - an analysis identifying research priorities and data requirements for resolving
  the red algal tree of life [@verbruggen2010data]
  - a review for ecologists about phylogenetic tools [@roquet2013building]
  - a study constructing a dated seed plant phylogeny using pyPHLAWD [@smith2018constructing]
  - a study presenting an assembly and alignment free method for phylogenetic reconstruction
    using genomic data, that aims to be incorporated in a tool as phylota some day [@fan2015assembly]
  - nexml format presentation [@vos2012nexml] - cites phylota as a tool that uses
    stored phyloinformatic data that could benefit from adopting nexml, to increase
    interoperability
  - a study of fruit evolution, analysing a previously published phylogeny of 8911
    tips of the Campanulidae, constructed with PHLAWD [@beaulieu2013fruit]
  - a study of Southeast Asia plant biodiversity inventory [@webb2010biodiversity] -
    cites phylota as a tool that would allow rapid phylogentic placing of newly
    discovered species, and generation of phylogenetically informed guides for field
    identification.

1. When the software was actually used to construct (partially or in full) a DNA
data set to be used for phylogenetic reconstruction:
  - A 1000 tip phylogeny of the family of the nightshades [@sarkinen2013solanaceae]
  - A 56 tip phylogeny of crustacean zooplancton [@helmus2010communities] -- ecological study
  - A 63 tip phylogeny of the Salmonidae family [@crete2012salmonidae]
  - A 321 tip phylogeny of Testudines [@thomson2010sparse]
  - A 69 taxa phylogeny of the family Cyprinodontidae of the pupfish [@martin2011trophic]
  - A 2,957 taxa phylogeny of the class Moniloformopses of living ferns [@lehtonen2011towards]
  - A 2,573 species phylogeny of the Papilionoidea [@hardy2014specialization]
  - A 23 taxa phylogeny of the California flora [@anacker2011origins]
  - Phylogenies of 6 different clades of flowering plants representing an independent
  evolutionary origin of extrafloral nectaries: *Byttneria* (Malvaceae), *Pleopeltis* (Polypodiaceae),
  *Polygoneae* (Polygoneaceae), *Senna* (Fabaceae), *Turnera* (Passifloraceae), and *Viburnum*
  (Adoxaceae) [@weber2014defense].
  - To supplement DNA data sets of various pre-existing mammalian phylogenetic trees
  sampled at different taxonomic levels [@faurby2015species]
  - A 900 species tree of muroid rodents, Muroidea [@steppan2017muroid], where 300
    species were newly added by the study and the rest obtained using phylota.
  - A 95 taxa phylogeny of Gymnosperms, focused on Ephedra, Gnetales [@ickert2009fossil]
  - A 1061 genera phylogeny of the Oscine birds [@selvatti2015paleogene]
  - A 268 species phylogeny of sharks, representing all orders and 32 families [@sorenson2014effect]

1. When the website was used to identify sequences and markers available in
GenBank for a particular group. In this cases, the dataset mining was either performed
with other tools, or not performed at all and just used for discussion:
  - A 812 tips phylogeny of the Order Chiroptera [@shi2015speciation] -- dataset
  constructed with PHLAWD
  - A 1276 tips phylogeny of the Fabaceae [@legume2013legume] -- dataset constructed
  by hand (I think??)
  - A review of dated phylogenies of fire-prone tropical savanna species from Brazil
  [@simon2012cerrado] -- just for discussion of the lack of markers available for
  these species on GenBank
  - A review of the phylogeetic sof the Apicomplexa, a parasitic phylum on unicellular
  protists [@morrison2009apicomplexa].
  - Three data sets from phylota (the suborder Pleurodira of side-necked turtles;
    the family Cactaceae of cacti; and the Amorpheae, a clade of legumes) were used
    to demonstrate and exemplify phylogenetic decisiveness [@sanderson2010phylogenomics]

1. Sometimes, it was cited by mistake:
    - In this 630 tip phylogeny of the Caryophyllaceae study [@greenberg2011caryophyllaceae] it might have been originally
    cited as an example of large phylogenies that reflect well supported relationships
    from previous smaller phylogenies. However, it was removed from the text but
    not from the final list of references. The DNA data set was constructed by hand most probably.