### 2020-06-24 submission_ams/submission_ams.Rmd

---
journal: jhm
layout: draft
title: Title here
author1: Andrew N. Other
author2: Fred T. Secondauthor
currentaddress: "Current address: Some other place, Germany"
affiliation: "American Meteorological Society,Boston, Massachusetts"
exauthors: 
  - name: Ping Lu
    exaffiliation: Princeton University
    correspondingauthor: "American Meteorological Society, 45 Beacon St., Boston, MA 02108."
    email: \email{groupleader@unknown.uu}
  - name: Miao Yu
    exaffiliation: University of Waterloo
    currentaddress: "Current address: Some other place, Canada"
abstract: |
  some abstract
bibliography: ../paper.bib
csl: american-meteorological-society.csl
output: rticles::ams_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r body, child="../2-body.Rmd"}
```