### 2020-06-24 submission_sysbio/submission-sysbio.Rmd

---
title: "Systematic Biology regular submission"
subtitle: "Submission and Formatting Instructions"
author:
  - name: First Author
    affiliation:
    - Affiliation 1
    - Affiliation 2
    email:  email@domain
  - name: Second Author
    affiliation:
    - Affiliation 1
    - Affiliation 2
    email:  email@domain
  - name: Third Author
    affiliation:
    - Affiliation 1
    - Affiliation 2
    email:  email@domain
abstract: |
  This document provides a basic template for the 2-page extended abstract and some submission guidelines. Including an abstract is optional, as the document itself is an extended abstract. If included, an abstract should be a single paragraph.
keywords: "Optional, Yet Potentially, Helpful"
output:
  bookdown::pdf_document2:
    template: sysbio-regular-template.tex
    keep_tex: yes
bibliography: paper.bib # Replace with your .bib file
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r body, child="2-body.Rmd"}
```