### 2020-06-24 _TODO.Rmd
methods specifications

---
title: Daily goals
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ilex}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(
  global.par = TRUE
  )
```

# next:

- [ ] Do not cite Felsenstein 1985 for bootstrap, just the call that we use for RAxML:
rax_ex, "-T", "{}".format(self.config.num_threads),
                                 "-m", "GTRCAT",
                                 "-s", alignment,
                                 "-p", "1", "-b", "1", "-#", str(num_reps),
                                 "-n", "BOOT{}".format(self.date)
- [ ] Finding a tree to update
  - [ ]  Using the curator website. Mention that there is a differnce between focal clade and ingroup. Physcraper does not take into account the focal clade from Open Tree. Because that is arbitrarily assigned by a curator. The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.
  - [ ] The link between the ingroup and the search taxon.

- [ ] How does the blast cycles results compare to other tools?

- [ ] Intro: why python? Introduce the taxon name space and the possibility of cross conectivity between platforms.

- [ ] How to create taxon name spaces for new trees so we can perform a comparison?

- [ ] How does pyphlawd, or phylota do the exemplar per species choosing for the baited analyses?

- [ ] bootsrap: is it classic [@felsenstein1985confidence] or transfer bootstrap expectation [@lemoine2018renewing]

- [ ] check the all-vs-all blast method, where does it come from?

- [ ] add phylotaR to the table of pipelines for mining
- [ ] check paper: A System for Phylogenetic Analyses over Alignments of Next Generation Sequence Data

- [ ] order of trees for synthesis https://files.opentreeoflife.org/synthesis/opentree12.3/opentree12.3/phylo_input/study_tree_pairs.txt

- [ ] comment on things needed for automatization of tree comparison
  - [ ] validating taxon names with a reference taxonomy
  - [ ] taxon sampling statistics
  - [ ] finding trees in a collection
  - [ ] pruning source trees to contain only taxa of interest

- [ ] the case for taxon name standardization
  - [ ] cite studies that have pointed out before that it is a necessary step for effective automated tree comparison @page2005taxonomic; @ranwez2009phyloexplorer; @stoltzfus2013phylotastic; treebase

- [ ] statistics on taxon coverage
  - [ ] dataset sparseness
  - [ ] taxonomic coverage of the datasets
  - [ ] taxonomic sampling among groups
  - [ ] detect under-represented taxa

- [ ] finding trees to update:
  - [ ] from a collection? NOT NOW
  - [ ] containing taxa from a specific taxon
  - [ ] trees from a study

- [ ] pruning trees

- [ ]  In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT for that single taxon would be determined as ***HOW?***
  - [ ] automatically, it will return the same taxon, even if its a species.
  - [ ] all tips in original tree are mapped to ott ids
  - [ ] sequences in alignment that are not mapped to ott ids will be kept in the alignment but not blast. Would only happen if somebody decides not to map a tip bc it is no longer the taxon they tought based on the new phylogenetic relationships.

# Done!

- [ ] determining the root
  - [ ] write something about the need to set the root to be able to do comparisons
  - [ ] say that utis hard to set it up automatically and that solutions are being sought

- [x] add bootsrap to methods

- [x] add BLAST command line tools to methods [@camacho2009blast]

- [X] explain how we root the updated trees

- [X] add blast citation!

- [x] add merced cluster nsf grant to acknowledgments
- [X] add genbank citations
  - [x] refer to it as a nucleotode database when decsribing the methods, bc it could be any molecular database, not just GenBank.

- [x] the ingroup
  - [x] how is it determined
  - [x] how is it relevant for blast results