### 2020-06-24 render2-submission-mee.Rmd

---
output:
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
geometry: margin = 1in
bibliography: paper.bib
csl: submission-mee/methods-in-ecology-and-evolution.csl
link-citations: yes
linestretch: 2
header-includes:
    - \usepackage{caption}
    - \usepackage{blindtext}
    - \usepackage[left]{lineno}
    - \linenumbers
params:
  author1: Luna L. Sanchez Reyes
  affiliation: School of Natural Sciences, University of California, Merced
---

`r params$affiliations`
Manuscript prepared for submission to Methods in Ecology and Evolution
Article type: Application

\newpage
# Summary

``` {r child = '1-abstract.Rmd'}
```

**Keywords**: phylogeny,

\newpage

``` {r child = '2-body.Rmd'}
```

\newpage
\begin{center}
\textsc{References}
\end{center}