### 2020-06-29 2.2-methods-ideas.Rmd
methods extras

# What does physcraper do?

Takes an alignment of any number of taxa.

Finds the mrca from the opentree synthetic tree, so it is not a taxonomic mrca, but a phylogenetic one (whenever there is phylogenetic information).

Blast sequences from the alignment, within the mrca context.

Blast can be done directly on NCBI (very slow, as all blast searches are) or on a copy of the NCBI database installed on a server (way faster).

Blasted sequences come back with NCBI taxonomy ids, so we need to match them to ott ids (using an api call).

Why not use phlawd or supersmart?

They are not integrated to the open tree workflow
physcraper allows tagging