### 2020-07-01 _TODO.Rmd
cleaning todos

---
title: Daily goals
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ilex}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(
  global.par = TRUE
  )
```

# next:

- [ ] hammer home the message that gene alignments are the gold standard for phylogenetics in many ways:
  - [ ] genomic data, uces, radseqs are not comparable to other data
  - [ ] homology in genomic data is not fully known (assembly based on similarity)
  - [ ] genomic is incorporating tests to try to curate alignments better
  - [ ] but data from gene sequences can only be used in a gene alignment context, not genomic data

- [ ] ilex remote example started on Sun Jun 28 at 20:07:32 PDT 2020 and finished on Mon Jun 29 at 06:33:57 PDT 2020

- [ ] on methods, bootstrap summary part cite:
 SumTrees: Phylogenetic Tree Summarization Version 4.4.0, from DendroPy 4.4.0 (main-bc20409)

- [ ] intro: challenge in connecting different parts of the database with other databases that might refer to the same thing. For example taxon ids represent own challenge in this sense, but it has been solved by the otol taxonomy project. Othe rchallenge is the study ids. This is not fully resolved yet by otol, but only partially by linking to the alignment repo.
- [ ] intro: programatic access.

- [ ] Finding a tree to update
  - [ ]  Using the curator website. Mention that there is a differnce between focal clade and ingroup. Physcraper does not take into account the focal clade from Open Tree. Because that is arbitrarily assigned by a curator. The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.
  - [ ] The link between the ingroup and the search taxon.

- [ ] How does the blast cycles results compare to other tools?

- [ ] Intro: why python? Introduce the taxon name space and the possibility of cross conectivity between platforms.

- [ ] How to create taxon name spaces for new trees so we can perform a comparison?

- [ ] How does pyphlawd, or phylota do the exemplar per species choosing for the baited analyses?

- [ ] check the all-vs-all blast method, where does it come from?

- [ ] add phylotaR to the table of pipelines for mining
- [ ] check paper: A System for Phylogenetic Analyses over Alignments of Next Generation Sequence Data

- [ ] order of trees for synthesis https://files.opentreeoflife.org/synthesis/opentree12.3/opentree12.3/phylo_input/study_tree_pairs.txt

- [ ] comment on things needed for automatization of tree comparison
  - [ ] validating taxon names with a reference taxonomy
  - [ ] taxon sampling statistics
  - [ ] finding trees in a collection
  - [ ] pruning source trees to contain only taxa of interest

- [ ] the case for taxon name standardization
  - [ ] cite studies that have pointed out before that it is a necessary step for effective automated tree comparison @page2005taxonomic; @ranwez2009phyloexplorer; @stoltzfus2013phylotastic; treebase

- [ ] statistics on taxon coverage
  - [ ] dataset sparseness
  - [ ] taxonomic coverage of the datasets
  - [ ] taxonomic sampling among groups
  - [ ] detect under-represented taxa

- [ ] finding trees to update:
  - [ ] from a collection? NOT NOW
  - [ ] containing taxa from a specific taxon
  - [ ] trees from a study

- [ ] pruning trees

- [ ]  In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT for that single taxon would be determined as ***HOW?***
  - [ ] automatically, it will return the same taxon, even if its a species.
  - [ ] all tips in original tree are mapped to ott ids
  - [ ] sequences in alignment that are not mapped to ott ids will be kept in the alignment but not blast. Would only happen if somebody decides not to map a tip bc it is no longer the taxon they tought based on the new phylogenetic relationships.

# Done!

- [ ] determining the root
  - [ ] write something about the need to set the root to be able to do comparisons
  - [ ] say that utis hard to set it up automatically and that solutions are being sought

- [x] add bootsrap to methods

- [x] add BLAST command line tools to methods [@camacho2009blast]

- [X] explain how we root the updated trees

- [X] add blast citation!

- [x] add merced cluster nsf grant to acknowledgments
- [X] add genbank citations
  - [x] refer to it as a nucleotode database when decsribing the methods, bc it could be any molecular database, not just GenBank.

- [x] the ingroup
  - [x] how is it determined
  - [x] how is it relevant for blast results

- [x] Do not cite Felsenstein 1985 for bootstrap, just explain it while explaining each argument from the call that we use for RAxML:
  rax_ex, "-T", "{}".format(self.config.num_threads),
                                   "-m", "GTRCAT",
                                   "-s", alignment,
                                   "-p", "1", "-b", "1", "-#", str(num_reps),
                                   "-n", "BOOT{}".format(self.date)

- [x] bootsrap: is it classic [@felsenstein1985confidence] or transfer bootstrap expectation [@lemoine2018renewing]
  - probably classic, but we are not citing, just explaining, see point above.