### 2020-07-02 2.2-methods-ideas.Rmd

# What does physcraper do?

Takes an alignment of any number of taxa.

Finds the mrca from the opentree synthetic tree, so it is not a taxonomic mrca, but a phylogenetic one (whenever there is phylogenetic information).

Blast sequences from the alignment, within the mrca context.

Blast can be done directly on NCBI (very slow, as all blast searches are) or on a copy of the NCBI database installed on a server (way faster).

Blasted sequences come back with NCBI taxonomy ids, so we need to match them to ott ids (using an api call).

Why not use phlawd or supersmart?

They are not integrated to the open tree workflow
physcraper allows tagging

Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.

***- One way to compare two independent phylogenetic trees is to compare them both to
the synthetic OToL and then measure how well they do against each other***

- Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon. This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
- Conflict information can only be generated in the context of the whole Open Tree of Life. Otherwise, it is not really possible to get conflict data.