### 2020-08-31 2-body.Rmd
word limit

---
title: 'Physcraper: Body'
author:
  - Luna L. Sanchez Reyes
  - Martha Kandziora
  - Emily Jane McTavish
  - University of California, Merced
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: paper.bib
csl: apa.csl
output:
  redoc::redoc
---

# Introduction

Phylogenetic estimates of evolutionary relationships capture the shared history of living and extinct organisms and provide key context for all our biological observations.
Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate. Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases. Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using phylogenetic knowledge from public biological databases to update a starting tree and single locus alignments.

Automatic integration of data in a phylogenetic context is challenged by the prevalence of taxonomic idiosyncrasies across biological databases. Taxonomic name standardization represents a step toward tackling this issue, but requires a unified system. The Open Tree of Life project (OpenTree)
<!-- [project](https://opentreeoflife.github.io/)  -->
has constructed a comprehensive and digitally-available tree of life by synthesizing published phylogenies along with taxonomic data. Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment. To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], the World Register of Marine Species (WoRMS), etc. The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.

Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving].
Genomics has, and will continue to, revolutionize phylogenetic inference.
Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
<!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have already generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.

More than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences. These sequences were gauged to have the potential to resolve phylogenetic relationships of most (`r round(236023/240708, digits=4)*100`%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota]. Assembling a DNA alignment from such a massive database can be done "by hand", but it requires huge amounts of time and it is mostly a non-reproducible approach. Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
<!-- *maybe a box of all pipelines?* -->
However, even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].

A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple]. The TreeBASE database [@piel2009treebase] hosts about 8, 200 alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a public source of valuable expert knowledge. Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life. The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated framework linking trees with the alignment that generated them.

Physcraper relies on programmatic access protocols (API's) to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses. Physcraper also provides new types of access to various OpenTree tools for comparison of existing phylogenetic hypotheses with newly generated ones.
Physcraper is coded as a Python pipeline that focuses on data interoperability, by using the standardized taxonomy as a way to link taxon data from different databases. This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, opening the possibility for a variety of comparative downstream analyses.

# The Physcraper framework

```{r framework, echo=FALSE, fig.cap="The Physcraper framework consists of 4 steps (see text). The software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, function usage descriptions, examples and tutorials.", out.width = '85%', fig.retina= 2, fig.align = "center"}
# knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
knitr::include_graphics("docs/figs/schematic.pdf")
```

The general Physcraper framework is depicted in Figure \@ref(fig:framework). It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.

## The inputs: a tree and an alignment

Taxon names in the input tree must be standardized or "mapped" to the unified OpenTree taxonomy [@rees2017automated] using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool. Users can choose from among the 2, 950 mapped trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase; @vos2012nexml].

The input alignment should be a single locus alignment that was used in part or in whole, to generate the input tree. If the alignment associated to the input tree is stred in TreeBASE, Physcraper retrieves it automatically. Alignments stored in any other repository, or constituting personal data have to be downloaded by the user.

Physcraper processes the input tree to match taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa. Technically, just one taxon name and its corresponding sequence in the alignment are needed to continue the algorithm.

## DNA sequence search and filtering

The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped], either on the GenBank remote database or in a GenBank local database set up by the user. It is constrained to a taxonomic group in the NCBI taxonomy, defined as the "search taxon". Users can arbitrarily define a search taxon that is either a more or a less inclusive
clade relative to the ingroup of the input tree.
Otherwise, the search taxon is automatically identified using the OpenTree [API](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated], as the Most Recent Common Ancestral Taxon (MRCAT) of the ingroup taxa in the input tree, i.e., the MRCA that is also a named clade in the NCBI taxonomy (Fig. \@ref(fig:framework)).
  <!-- FIGURE RECOMMENDED: Figure \@ref(fig:search)  -->
The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.

BLAST command line tools `blastn` function [@camacho2009blast] is used for local database sequence searches.
For remote database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL). This is useful when a user lacks access to the computer capacity needed to setup a local database, but has access to an institutional server.
Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank within the search taxon. GenBank sequences with match scores smaller than the e-value cutoff (default to 0.00001) are downloaded into a local library.

Downlodaded sequences are not further considered in the analysis if they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80% and 120%, respectively); or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OTT taxonomy or the NCBI taxonomy.
If there are several sequences per taxon, an arbitrary number of 5 (can be modified by the user) sequences per taxon are chosen at random.

Reverse, complement, and reverse-complement sequences are identified and translated using BioPython internal functions [@cock2009biopython].
Iterative cycles of BLAST searches can be performed, by blasting the new sequences until no new ones are found. By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.

## New DNA sequence alignment

By default, Physcraper uses MUSCLE [@edgar2004muscle] to perform DNA sequence alignments in a two step process. First, all new sequences are aligned using the default MUSCLE options.

Second, a MUSCLE profile alignment is performed, in which the original alignment is used as a template to align the new sequences. This ensures that the final alignment
follows the homology criteria established by the original alignment.
The final alignment is not further processed by Physcraper. It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [e.g., GBlocks; @castresana2000selection; @castresana2002gblocks].
Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.

## Tree reconstruction and comparison

A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with default settings (GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm). Only the number of bootsrap replicates can be defined by the user.
By default, the original tree is used as a starting tree for the ML searches. Alternatively, the original tree can be set as full topological constraint, or simply be ignored for the ML searches.
Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].

Physcraper's final result is an updated phylogenetic hypothesis for the search taxon based on the data in the input alignment.
There are two ways to compare the updated tree with the original tree implemented in Physcraper. First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions [@sukumaran2010dendropy].
Second, a conflict analysis is performed. This is a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict Application Programming Interface [@redelings2017supertree].
For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined. A suggested default rooting based on OpenTree's taxonomy is implemented for now. This approach uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree. However, if the updated tree changes expectations from taxonomy, the root may no longer be appropriate. Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet. The best way right now is for users to define the outgroup directly on the updated tree, so trees are accurately rooted.
<!-- *It would be a nice addition to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon* -->

# Example: The hollies

A user is interested in the phylogenetic relationships within the genus *Ilex*. Commonly known as "hollies", the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.

An online literature review in June 2020 (google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].
@gottlieb2005molecular made original tree and alignment data available in [TreeBASE](https://treebase.org/treebase-web/search/study/summary.html?id=1091). The "Gottlieb2005" tree sampling 41 species was added to the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home) and its information has been integrated into [OpenTree's synthetic tree](https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa).

The most recent *Ilex* tree from @yao2020phylogeny, has been made available in the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/view/ot_1984) and in the [DRYAD repository](https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x). The "Yao2020" tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.

A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
Figure \@ref(fig:results) shows results from the Physcraper analysis of an alignment of the internal transcribed spacer DNA region (ITS) from @gottlieb2005molecular. Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
<!-- **MTH: you probably need some details about the hardware, given the fact that you are discussing running times** -->
The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional *Ilex* taxa. The best RaxML tree is 99% resolved, with 25% of nodes with bootstrap support < 0.1 and 48% nodes with bootstrap support > 0.75.
A large portion of internal branches are negligibly small, with 30 branches < 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75 (Fig. \@ref(fig:results)).
For comparison, the Yao2020 tree also contains all 41 distinct taxa from the original Gottlieb2005 study,
<!-- has less new tips (134), but  -->
and contributes phylogenetic data to 134 additional *Ilex* taxa, from which
`r 135-68` are also in the Physcraper updated Gottlieb2005 tree. While @yao2020phylogeny also used ITS as a marker, their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis. However, Physcraper was able to incorporate 18 taxa that were not in Yao2020.
<!-- **Any statements about relationships or anything? Are there multiple seqs for these taxa?**
add 1 sentence saying that it is odd that we are getting differnet relationships given what is likely to be the same data
but we can't compare the exact seqs or alignemnt, bc they are not available.
 -->
This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication, but seems to be a "manual" process.
<!-- More Physcraper examples at a range of taxonomic scales are available in the Physcraper documentation online. -->
<!-- The alignments have THIS MANY columns and THIS AMOUNT of missing data. -->
<!-- RF DISTANCE INTERPRETATION -->
<!-- ADD ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE ?-->

<!-- ```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
wget -O docs/figs/ilex-plot-tips-and-branches-1.png https://raw.githubusercontent.com/McTavishLab/physcraperex/master/docs/articles/ilex_files/figure-html/plot-tips-and-branches-1.png
``` -->

```{r results, echo=FALSE, fig.cap="A) Phylogenetic tree obtained by updating the Gottlieb et al. 2005 tree in B) using Physcraper.", out.width = '100%', fig.retina= 2}

# knitr::include_graphics("docs/figs/ilex-plot-tips-and-branches-1.png")
knitr::include_graphics("docs/figs/ilex-results.png")

```

\newpage

*Figure \@ref(fig:results) caption continued*: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively. Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section "DNA sequence search and filtering", aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps. B-D show results of the conflict analysis performed with OpenTree tools.

# Discussion

Data repositories designed to preserve and democratize access to biological data
constitute an essential resource for evolutionary estimation. While data keep accumulating,
our ability to reanalyze and incorporate new data with available knowledge is being challenged.

Various phylogenetic pipelines have been designed to make evolutionary sense of the vast amount of public molecular data available in different ways.
Pipelines like Phylota [@sanderson2008phylota], PyPHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward]
focus mainly on generating full trees *de novo* (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates) for regions of the Tree of
Life that have no phylogenetic assessment yet in published studies.
While Physcraper does not generate phylogenies *de novo* in a traditional sense,
it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
phylogenetic knowledge available in tree repositories to facilitate phylogenetic placement of public molecular data.
The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
pre-existing alignments to incorporate public molecular data into phylogenies. Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
<!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->
Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving]. However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g. ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], SVD Quartets [@chifman2014quartet]).

As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
It can also link to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is contributing to the Encyclopedia of Life main goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].

As such, Physcraper can be used to rapidly (in a matter of hours)
solve challenges overarching both fields of ecology and evolution, such as
placing newly discovered species phylogenetically [@webb2010biodiversity],
obtaining trees for ecophylogenetic studies [@helmus2012phylogenetic],
systematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
and generating custom species trees for ecological and evolutionary downstream analyses [@stoltzfus2013phylotastic].

Data repositories hold more information than meets the eye.
Besides the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
Usually, initial ideas about the data are changed by new analyses.
Ideally, new understanding of the data can be continually registered on databases
to consistently expose newcomers to the most up to date knowledge about the data.

# Acknowledgements

Research was supported by the grant "Sustaining the Open Tree of Life", National Science Foundation ABI No. 1759838, and ABI No. 1759846.
Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No. ACI-1429783.

We thank the members of the OpenTree development team and the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on this manuscript.

The authors have no conflict of interest to declare.

# Authors' Contributions

LLSR wrote the manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
All authors contributed to the manuscript and gave final approval for publication.

# Data Archiving

Physcraper source code available at https://github.com/McTavishLab/physcraper

Documentation available at https://physcraper.readthedocs.io/en/latest/index.html

Illustrated examples available at https://github.com/McTavishLab/physcraperex

This is a reproducible manuscript available at https://github.com/McTavishLab/physcraper_ms

\newpage
# References