### 2020-11-18 sources/2pdf-bmc.Rmd
try render bmc format pdf from rmd

---
output:
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
geometry: margin = 1in
bibliography: paper.bib
csl: bmc-bioinformatics.csl
link-citations: yes
linestretch: 2
header-includes:
    - \usepackage{color}
    - \usepackage{hyperref}
    - \usepackage{caption}
    - \usepackage{blindtext}
    - \usepackage{url}
    - \usepackage[left]{lineno}
    - \linenumbers
title: 'Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life'
params:
  running_title: Updating gene trees with the Open Tree of Life
  twitter: 'Physcraper: a tool to update published phylogenies and link taxa to biological databases using the Open Tree of Life'
  author1: Luna L. Sanchez Reyes
  email1: sanchez.reyes.luna@gmail.com
  author2: Martha Kandziora
  email2: kandziom@natur.cuni.cz
  author3: Emily Jane McTavish
  email3: ejmctavish@ucmerced.edu
  affiliation: School of Natural Sciences, University of California, Merced
  affiliation2: Department of Botany, Faculty of Science, Charles University, Prague, Czech Republic
  correspondence_address: Science and Engineering Building 1, University of California, Merced, 5200 N. Lake Rd, Merced CA 95343
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  eval = TRUE
)
library(wordcountaddin)
word_count_body <- wordcountaddin::word_count(filename="2-body.Rmd")
word_count_references <- 873
```
**1. `r params$author1`**

`r params$affiliation`

email: <`r params$email1`>

**2. `r params$author2`**116*0.57

`r params$affiliation`

`r params$affiliation2`

email: <`r params$email2`>

**3. `r params$author3`**

`r params$affiliation`

email: <`r params$email3`>

**Correspondence address**: `r params$correspondence_address`

**Correspondence email**: <`r params$email1`>, <`r params$email3`>

**Running title**: `r params$running_title`

**Word count**: `r word_count_body + word_count_references`

**Manuscript prepared for submission to Methods in Ecology and Evolution**

**Article type**: Application

\newpage

\begingroup\Large
**Abstract**
\endgroup

``` {r child = '1-abstract.Rmd'}
```

**Keywords**: gene tree, interoperability, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, updated alignment

\newpage

``` {r child = '2-body.Rmd'}
```