### 2021-01-06 temp.Rmd
intro

While genomics has, and will continue to, revolutionize phylogenetic inference,
the diversity of alternative genomic sequencing approaches implemented produce
largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
creating challenges for phylogenetic reconstruction.
Phylogenomics addresses this problem by focusing on targeted capture of informative
regions \cite{andermann2020guide}. Yet, it is expensive. Few labs around the world can afford
to implement it for all their study groups, or at all. Creating a wealth gap for
science, while free information is still available to provide new knowledge.
Genomics does not contribute yet to democratization of phylogenetic research \cite{li2017enabling}.

%% Is genomic data on genbank available for a smaller percentage of all life, compared to
%% single loci in the same database?

Public biological databases constitute an amazing resource for evolutionary
studies. Connecting existing phylogenies to molecular data that has never been
incorporated into any phylogenetic estimate, geographical location, fossils, and
other data in a reproducible and continuous manner is possible by establishing a
data interoperability framework for biological databases. Here, we introduce
Physcraper, a tool that automates biological database connections
with the main goal of building upon homology
hypotheses that taxon specialists have assessed and deemed appropriate for a
specific phylogenetic scope to update a starting tree and single locus alignments
with public DNA data.