### 2021-01-07 temp.Rmd
intro

While genomics has, and will continue to, revolutionize phylogenetic inference,
the diversity of alternative genomic sequencing approaches implemented produce
largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
creating challenges for phylogenetic reconstruction.
Phylogenomics addresses this problem by focusing on targeted capture of informative
regions \cite{andermann2020guide}. Yet, it is expensive. Few labs around the world can afford
to implement it for all their study groups, or at all. Creating a wealth gap for
science, while free information is still available to provide new knowledge.
Genomics does not contribute yet to democratization of phylogenetic research \cite{li2017enabling}.

%% Is genomic data on genbank available for a smaller percentage of all life, compared to
%% single loci in the same database?

\cite{pyron2011phylogeny} PHLAWD, Colubroidea, 334 citations

Large genomic dataset limit the complexity of evolutionary models that can be applied.
Single locus allows for more complex evolutionary models, and therefore might
provide better evolutionary estimates \cite{}.

Single locus super-matrix have provided very good phylogenegic inferences, sometimes even better than genomic based estimates. SL also, allow for construction of large phylogenies and

Line 78- Add extra sentence about advantages of Physcraper's implementation
Physcraper has speciifc advantages over existing pipelines.