### 2021-01-07 temp.Rmd
intro and discussion

While Physcraper does not generate phylogenies \textit{de novo} in a traditional sense,
it successfully generates new phylogenetic knowledge, revealing the importance of
open science in facilitating phylogenetic placement of public molecular data and
accelerating enrichment and updating of phylogenetic relationships in any region
of the tree of life.

Genomics is expensive. Few labs around the world can afford
to implement it for all their study groups, or at all. Creating a wealth gap for
science, while free information is still available to provide new knowledge.
Genomics does not contribute yet to democratization of phylogenetic research \cite{li2017enabling}.

%% Is genomic data on genbank available for a smaller percentage of all life, compared to
%% single loci in the same database?

%% While taxonomic trees are useful, trees inferred from evolutionary estimates are
preferred.

Notable examples of phylogenies constructed using multi-locus, database mined, super matrix approach:
Magallon, Soltis, examples with fish, bats,
\cite{magallon2015metacalibrated} mining by hand - flowering plant families - has been cited 561 times
\cite{smith2018constructing} 38k clades of plants with seeds - cited 176 times
\cite{jetz2012global} 6k bird species - cited
\cite{rabosky2018inverse} mining with phlawd - cited 202 times

These studies are widely cited and immensely relevant for the biological community.

Some of these multilocus phylogenetic studies have implemented pipelines such as PHLAWD,
other studies have curated their matrices by hand, which is largely non-reproducible.

Make emphasis on these really great studies (and pipelines) that do not rely on
genomic data, and that there is a lot of single locus data that can be analysed.

\cite{pyron2011phylogeny} PHLAWD, Colubroidea, 334 citations

Large genomic dataset limit the complexity of evolutionary models that can be applied.
Single locus allows for more complex evolutionary models, and therefore might
provide better evolutionary estimates \cite{}.

Single locus super-matrix have provided very good phylogenegic inferences, sometimes even better than genomic based estimates. SL also, allow for construction of large phylogenies and

Line 78- Add extra sentence about advantages of Physcraper's implementation
Physcraper has speciifc advantages over existing pipelines.